---
title: "Data screening"
author: "Tristan Mahr"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include = FALSE}
library("knitr")
opts_chunk$set(
  cache.path = "assets/cache/01-",
  fig.path = "assets/figure/01-",
  warning = FALSE,
  collapse = TRUE, 
  comment = "#>", 
  message = FALSE,
  fig.width = 8,
  fig.asp = 0.618,
  dpi = 300,
  out.width = "80%")
options(width = 100)
```

This script screens the eyetracking data by removing unreliable trials and
unreliable blocks of trials. It preserves the pairs of matched children. (If a
child is removed for have no reliable eyetracking data, their match is also
removed.) Counts of trials and children at each step of screening are printed
too.

## Set up

Load the data.

```{r}
library(dplyr, warn.conflicts = FALSE)
df_child <- readr::read_csv("data-raw/test-scores.csv") %>% 
  select(Group, Matching_PairNumber, 
         ChildID, ChildStudyID, Study, ResearchID, Female)
df_blocks <- readr::read_csv("data-raw/blocks.csv")
df_looks <- readr::read_csv("data-raw/looks.csv.gz")
df_trials <- readr::read_csv("data-raw/trials.csv")
```

In order to aggregate looks with my littlelisteners package, we need to define 
a response coding definition so it knows what's a target, what's offscreen, etc.

```{r}
library(littlelisteners)
def <- create_response_def(
  label = "LWL scheme",
  primary = "Target",
  others = "Distractor",
  elsewhere = "tracked", 
  missing = NA)
```

Set data-screening options.

```{r}
screening <- list(
  # determine amount of missing data between:
  min_time = 250,
  max_time = 1500,
  # remove trials with more than ... proportion of missing data
  max_na = .5,
  # blocks should have at least this many trials
  min_block = 12,
  # conditions should have at least this many trials
  min_condition = 6
)
```

## Remove blocks from the bad version of the experiment

The first version of the experiment ran too quickly; there wasn't much time 
between the prompt and the reinforcer phrase.

```{r}
blocks_to_drop <- df_blocks %>% 
  filter(Block_Version != "Standard") %>% 
  select(BlockID:Block_Version)

knitr::kable(blocks_to_drop)

looks_good_version <- df_looks %>% 
  anti_join(blocks_to_drop)
```


## Find unreliable trials

We identify trials with more than `r screening$max_na * 100`%
missing data during a trial window (here `r screening$min_time`--
`r screening$max_time` ms).

```{r}
trial_quality <- looks_good_version %>% 
  left_join(df_child) %>% 
  # Offset by 20 ms because the data binned into 50ms bins: I.e., the frame at
  # 285ms is part of the [285, 300, 315] ms bin, so that frame needs to part of
  # the data screening.
  filter(between(Time, screening$min_time - 20, screening$max_time + 20)) 

range(trial_quality$Time)

missing_data_per_trial <- trial_quality %>% 
  aggregate_looks(def, Group + Study + ResearchID + 
                    Matching_PairNumber + TrialID ~ GazeByImageAOI) %>% 
  select(Group:TrialID, PropNA) %>% 
  print()
```

Count the number of bad trials that need to be excluded.

```{r}
# Using UQ to unquote `screening$max_na` so that the created column is
# *not* named `PropNA > screening$max_na` but instead uses the value of
# `screening$max_na` in the column name.
missing_data_per_trial %>% 
  count(PropNA > UQ(screening$max_na)) %>% 
  rename(`Num Trials` = n) %>% 
  knitr::kable()

missing_data_per_trial %>% 
  count(Group,PropNA > UQ(screening$max_na)) %>%  
  rename(`Num Trials` = n) %>% 
  knitr::kable()
```

Remove the bad trials.

```{r}
bad_trials <- missing_data_per_trial %>% 
  filter(PropNA >= screening$max_na)

looks_clean_trials <- anti_join(looks_good_version, bad_trials)
```

## Find unreliable blocks

Count the number of trials leftover in each block.

```{r}
# Narrow down to one row per trial
trial_counts <- looks_clean_trials %>% 
  select(BlockID:ResearchID) %>% 
  distinct() %>% 
  # Count rows in each group
  count(BlockID, Study, ResearchID) %>% 
  arrange(n) %>% 
  rename(`Num Trials in Block` = n)
```

Identify blocks with fewer than `r screening$min_block` trials.

```{r}
blocks_to_drop <- trial_counts %>% 
  filter(`Num Trials in Block` < screening$min_block) 

knitr::kable(blocks_to_drop)
```

Remove the blocks.

```{r}
looks_clean_blocks <- looks_clean_trials %>% 
  anti_join(blocks_to_drop) 
```



## Enforce minimum number of trials per condition

Count the number of trials leftover in each condition.

```{r}
# Narrow down to one row per trial
condition_counts <- looks_clean_blocks %>% 
  left_join(df_trials %>% select(TrialID, Condition)) %>% 
  select(Study, ResearchID, TrialID, Condition) %>% 
  distinct() %>% 
  # Count rows in each group
  count(Condition, Study, ResearchID) %>% 
  # Make sure 0 trials get counted
  tidyr::complete(tidyr::nesting(Study, ResearchID), Condition, 
                  fill = list(n = 0)) %>% 
  arrange(n) %>% 
  rename(`Num Trials in Condition` = n) %>% 
  print()
```

We want to ensure that there are at least 
`r screening$min_condition` trials per condition.

```{r}
children_to_drop <- condition_counts %>% 
  filter(`Num Trials in Condition` < screening$min_condition) 

if (nrow(children_to_drop) == 0) {
  print("No children to remove")
} else {
  knitr::kable(children_to_drop)
}
```

Remove the children.

```{r}
looks_clean_conditions <- looks_clean_blocks %>% 
  anti_join(children_to_drop) 
```


## Remove unpaired children

Next, we need to exclude participants who no longer have a match.

```{r}
df_leftover <- looks_clean_conditions %>% 
  distinct(Study, ResearchID) %>% 
  left_join(df_child)

df_leftover %>% 
  count(Group) %>% 
  ungroup() %>% 
  rename(`Num Children` = n)

df_singletons <- df_leftover %>% 
  count(Matching_PairNumber) %>% 
  ungroup() %>% 
  rename(NumChildrenInPair = n) %>% 
  filter(NumChildrenInPair == 1) %>% 
  print()

df_looks_clean <- looks_clean_conditions %>% 
  left_join(df_child %>% 
               select(Group, Matching_PairNumber, Study, ResearchID)) %>% 
  anti_join(df_singletons)
```

Now, there will be the same number of children in each group x task.

```{r}
df_looks_clean %>% 
  distinct(Group, Study, ResearchID) %>% 
  count(Group) %>% 
  ungroup() %>% 
  rename(`Num Children` = n) %>% 
  knitr::kable()

# Make sure there are 2 children in every matching pair
df_looks_clean %>% 
  distinct(Matching_PairNumber, Group, Study, ResearchID) %>% 
  count(Matching_PairNumber) %>% 
  ungroup() %>% 
  rename(NumChildrenInPair = n) %>% 
  filter(NumChildrenInPair != 2)
```








## Data screening counts

Count the number of children and trials at each stage in data screening.

```{r}
cleaning_progression <- list(
  `a. raw data` = df_looks,
  `b. remove bad experiment version` = looks_good_version,
  `c. drop bad trials` = looks_clean_trials, 
  `d. drop sparse blocks` = looks_clean_blocks,
  `e. drop children w sparse conditions` = looks_clean_conditions,
  `f. drop unpaired children` = df_looks_clean) %>% 
  bind_rows(.id = "Stage") %>% 
  select(-Group, -Matching_PairNumber) %>% 
  left_join(df_child) 

cleaning_progression %>% 
  group_by(Stage) %>% 
  summarise(
    `Num Pairs` = n_distinct(Matching_PairNumber),
    `Num Children` = n_distinct(ChildID),
    `Num Child-Study IDs` = n_distinct(ChildStudyID),
    `Num Blocks` = n_distinct(BlockID),
    `Num Trials` = n_distinct(TrialID)) %>% 
  knitr::kable()

cleaning_progression %>% 
  group_by(Stage, Group) %>% 
  summarise(
    `Num Children` = n_distinct(ChildID),
    `Num Child-Study IDs` = n_distinct(ChildStudyID),
    `Num Blocks` = n_distinct(BlockID),
    `Num Trials` = n_distinct(TrialID)) %>% 
  knitr::kable()
```


## Clean up

```{r}
df_looks_w_info <- df_looks_clean %>% 
  left_join(df_blocks) %>% 
  left_join(df_trials) %>% 
  left_join(df_child) %>% 
  select(ChildID, ChildStudyID, Group, Matching_PairNumber, Study, ResearchID, 
         Block_Basename, Block_Dialect, StimulusSet, Block_Version,
         TrialNo, Condition:TargetImage, Time:GazeByAOI)

df_looks_w_info %>% 
  distinct(Block_Dialect, StimulusSet, Block_Version)

df_looks_w_info <- df_looks_w_info %>% 
  select(-Block_Dialect, -Block_Version, -StimulusSet)
```

Save clean data.

```{r}
df_looks_w_info %>% 
  mutate(
     Group_Lab = Group %>% 
      factor(c("CochlearImplant", "NormalHearing"), 
             c("Cochlear implant", "Normal hearing"))) %>% 
  readr::write_csv("./data/screened.csv.gz")
```

Update participants data to only have children with eyetracking data.

```{r}
readr::read_csv(file.path(".", "data-raw", "test-scores.csv")) %>% 
  semi_join(df_looks_w_info) %>% 
  mutate(
    Group_Lab = Group %>% 
      factor(c("CochlearImplant", "NormalHearing"), 
             c("Cochlear implant", "Normal hearing"))) %>%
  readr::write_csv("./data/scores.csv")
```



***

```{r}
sessioninfo::session_info()
```


