---
title: "Consider the first image viewed"
author: "Tristan Mahr"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include = FALSE}
library("knitr")
opts_chunk$set(
  cache.path = "assets/cache/04-",
  fig.path = "assets/figure/04-",
  warning = FALSE,
  collapse = TRUE, 
  comment = "#>", 
  message = FALSE,
  fig.width = 8,
  fig.asp = 0.618,
  dpi = 300,
  out.width = "80%")
options(width = 100)
```

This script fits an exploratory growth curve model where looks are 
re-aggregated based on where the child fixated at the onset of the noun.

## Set up

```{r, results = 'hide'}
library(dplyr)
library(littlelisteners)
library(ggplot2)
library(lme4)

source("./plotting-helpers.R", encoding = "UTF8")
looks <- readr::read_csv("./data/bias.csv.gz") %>% 
  mutate(
    Cond_Lab = Condition %>% 
      factor(c("real", "MP", "nonsense"),
             c("Real word", "Mispronunciation", "Nonword")))
```

### Add orthogonal polynomials

```{r}
looks <- looks %>% 
  polypoly::poly_add_columns(Time, degree = 3, prefix = "ot") 
```

### Prep the datasets

For the models with vocabulary information, we need to remove any pairs of
children in which one of the children is missing an EVT score. We identify those 
pairs.

```{r}
no_vocab_pairs <- looks %>%
  distinct(ChildStudyID, Matching_PairNumber, EVT_Standard) %>%
  filter(is.na(EVT_Standard)) %>%
  select(Matching_PairNumber) %>%
  print()
```

Prepare a dataset for each condition.

```{r}
d_mp <- looks %>% 
  filter(Condition == "MP")

d_rw <- looks %>% 
  filter(Condition == "real")

d_ns <- looks %>% 
  filter(Condition == "nonsense")

d_mp_evt <- d_mp %>% 
  anti_join(no_vocab_pairs, by = "Matching_PairNumber")

d_rw_evt <- d_rw %>% 
  anti_join(no_vocab_pairs, by = "Matching_PairNumber")

d_ns_evt <- d_ns %>% 
  anti_join(no_vocab_pairs, by = "Matching_PairNumber")
```

## Group by initial image models

For these models we include an indicator for whether the child fixated on the
familiar or novel object during target noun onset. This predictor doubled the
number of (growth curve shape) parameters in our model and cuts the number of
observations in cell by more than half. (More than half because in some trials
the fixated at neither the familiar or novel object, so those trials have to be
excluded.) So take these with a grain of salt.

```{r no-evt-models, cache = TRUE}
glmer_controls <- glmerControl(
  optimizer = "bobyqa",
  optCtrl = list(maxfun = 2e5))

m_mp <- glmer(
  cbind(Target, Distractor) ~ 
    Bias_ImageAOI * Group * (ot1 + ot2 + ot3) + 
    (ot1 + ot2 + ot3 | ChildStudyID:Bias_ImageAOI),
  family = binomial,
  control = glmer_controls,
  data = d_mp)
summary(m_mp)

m_rw <- glmer(
  cbind(Target, Distractor) ~ 
    Bias_ImageAOI * Group * (ot1 + ot2 + ot3) + 
    (ot1 + ot2 + ot3 | ChildStudyID:Bias_ImageAOI),
  family = binomial,
  control = glmer_controls,
  data = d_rw)
summary(m_rw)

m_ns <- glmer(
  cbind(Target, Distractor) ~ 
    Bias_ImageAOI * Group * (ot1 + ot2 + ot3) + 
    (ot1 + ot2 + ot3 | ChildStudyID:Bias_ImageAOI),
  family = binomial,
  control = glmer_controls,
  data = d_ns)
summary(m_ns)
```


We can plot the growth curve fixed effects which describe how the average child
in each group x condition x initial-look performs.

```{r overall-fits, echo = FALSE}
scale_plasma <- viridis::scale_color_viridis(
  discrete = TRUE, 
  option = "plasma", 
  end = .7)

d_times <- d_ns %>% 
  distinct(Bias_ImageAOI, Time, ot1, ot2, ot3, Group) %>% 
  filter(!is.na(Bias_ImageAOI)) %>% 
  mutate(`First Look` = ifelse(Bias_ImageAOI == "Target", 
                               "Familiar object", "Novel object"))

condition_labels <- looks %>% 
  distinct(Condition, Cond_Lab, Group, Group_Lab)

models <- list(
  nonsense = m_ns,
  MP = m_mp,
  real = m_rw)

fitted <- map_predict(d_times, models, predict_fixef_resp) %>% 
  left_join(condition_labels, 
            by = c("model" = "Condition", "Group"))

only_bias_looks <- looks %>% 
  filter(!is.na(Bias_ImageAOI)) %>% 
  mutate(`First Look` = ifelse(Bias_ImageAOI == "Target", 
                               "Familiar object", "Novel object"))

ggplot(only_bias_looks) +
  theme_grey(base_size = 14) +
  aes(x = Time, color = Group_Lab, group = interaction(Group_Lab, Cond_Lab)) + 
  hline_chance() +
  stat_mean_se(mapping = aes(y = Prop), alpha = .4, shape = 1) + 
  geom_line(aes(y = fitted, linetype = Group_Lab, 
                group = interaction(Group_Lab, Cond_Lab)), data = fitted, 
            size = 1) +
  scale_plasma +  
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  labs(x = plot_text$x_time,
       y = plot_text$y_fits,
       color = "Group",
       linetype = "Group") + 
  legend_top() + 
  align_axis_right() + 
  facet_grid(`First Look` ~ Cond_Lab) + 
  ggtitle("Looks by image viewed during word onset")

for (x in unique(only_bias_looks$Cond_Lab)) {
  this_looks <- only_bias_looks %>% filter(Cond_Lab == x)
  this_fitted <- fitted %>% filter(Cond_Lab == x)
  p <- ggplot(this_looks) +
    theme_grey(base_size = 14) +
    aes(x = Time, color = Group_Lab, group = interaction(Group_Lab, Cond_Lab)) + 
    hline_chance() +
    stat_mean_se(mapping = aes(y = Prop), alpha = .4, shape = 1) + 
    geom_line(aes(y = fitted, linetype = Group_Lab,
                  group = interaction(Group_Lab, Cond_Lab)), data = this_fitted, 
              size = 1) +
    scale_plasma +  
    guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    labs(x = plot_text$x_time,
         y = plot_text$y_fits,
         color = "Group",
         linetype = "Group",
         caption = paste0("Condition: ", x)) + 
    legend_top() + 
    align_axis_right() + 
    facet_grid(. ~ `First Look`, labeller = label_both) 
    expand_limits(y = 0:1)
  print(p)
}
```


***

```{r}
sessioninfo::session_info()
```



